\textbf{Analyzing Machupo virus-receptor binding by molecular dynamics simulations}

### 2014-01-06 paper.tex
Changed pretty much every table and figure as well as all of the scripts.

The statistical package R was used for all further analysis and visualization. Each of the 50 independent trajectories per mutant produced a fairly noisy force curve. The force curves for each mutant were smoothed either by averaging over many replicates as in Figure~\ref{fig:force_curve} or by selecting a 50 frame interpolation window for averaging. The latter approach was used for statistical analysis. The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC). We used the pairwise.t.test function with the FDR p-value correction to test for significance. The ggplot \citep{ggplot} package was used to generate most of the figures.

@@

We proceeded to compare our SMD results to that of the standard dual topology FEP approach to calculate relative free energy differences. The correlation between the energetically rigorous FEP and our statistical approach is high. For all 11 complexes tested, the correlation between max force and FEP was $r=-0.767$ at $p=0.005857$ (Figure~\ref{fig:fep}), and the correlation between AUC and FEP was $r=--0.729$ at $p=0.011$. Because of this strong correlation, we refer exclusively to the SMD results for the remainder of this work.

@@

\clearpage
\begin{figure}[p]
\centerline{\includegraphics[width=6in]{figures/fep_mutants.pdf}}
\caption{\label{fig:fep}\title{Max force versus free energy perturbation.} Scatter plot of maximum force in SMD versus the relative free energy difference calculated by FEP for all 10 mutants tested plus the WT complex. The WT complex for FEP was simply set to 0.0. The correlation between the two is $r=-0.767$ with $p=0.005857$.}
\end{figure}

@@

\clearpage
\begin {table}[H]
\caption{\label{tab:summary_mutants}Summary of prior information available for each mutation tested. Observed $in~vivo$ refers to mutations that have been observed in rodent populations. Phenotype $in~vitro$ refers to the observed phenotype in $in~vitro$ viral entry assays.} 
\begin{center}
  \resizebox{11cm}{!} {
    \begin{tabular}{l c r r}
    \hline
      Mutation & Observed $in~vivo$ & Phenotype $in~vitro$ \\ \hline
            WT &                Yes &         Normal Entry \\
         N348A &                 No &                    - \\
         N348K &                Yes &     Diminished Entry \\
         N348W &                 No &                    - \\
         Y211A &                 No &        No Expression \\
         Y211D &                Yes &        No Expression \\
         Y211T &                 No &     Diminished Entry \\
        vR111A &                 No &     Diminished Entry \\
   N348A/Y211A &                 No &                    - \\
   N348W/Y211A &                 No &                    - \\
  vR111A/Y211A &                 No &                    - \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

\clearpage
\begin {table}[H]
\caption{\label{tab:summary_statistics}Summary statistics for each mutation tested. $\mu_\text{MAF}$ is the mean in piconewtons and $\sigma_\text{MAF}$ is the standard deviation of maximum applied force over all simulations. $\mu_\text{AUC}$ is the mean and $\sigma_\text{AUC}$ is the standard deviation of AUC over all simulations. $\Delta G$ is the free energy difference in kcal/mol calculated via FEP by the dual topology paradigm.}
\begin{center}
  \resizebox{15cm}{!} {
    \begin{tabular}{l c c c c c}
    \hline
      Mutation & $\mu_\text{MAF}$ (pN) & $\sigma_\text{MAF}$ & $\mu_\text{AUC}$ & $\sigma_\text{AUC}$ & $\Delta G$ (kcal/mol)\\ \hline
            WT &                772.00 &              154.13 &         198794.7 &            66765.64 &  0.000\\
         N348A &                799.33 &              154.36 &         202247.3 &            65175.78 & -2.149\\
         N348K &                731.36 &              151.01 &         179087.5 &            55799.49 & +3.184\\
         N348W &                732.13 &              131.21 &         192043.2 &            66916.36 & +3.033\\
         Y211A &                684.03 &              126.69 &         153517.7 &            40889.12 & +2.526\\ 
         Y211D &                875.44 &              116.46 &         234909.3 &            50494.11 & -2.760\\ 
         Y211T &                861.17 &              158.48 &         231078.7 &            97569.29 & +0.876\\
        vR111A &                754.39 &              103.25 &         185896.8 &            46408.74 & +0.467\\
   N348A/Y211A &                762.96 &              144.81 &         173059.2 &            59898.10 & +5.203\\
   N348W/Y211A &                617.21 &              163.33 &         147705.6 &            48548.61 & +8.207\\
  vR111A/Y211A &                780.71 &              135.65 &         189919.1 &            48868.16 & -2.441\\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

\clearpage
\begin {table}[H]
\caption{\label{tab:MAF_pairwise}Pairwise differences (row variable minus column variable) in mean maximum applied force. Bolded values are statistically significant at $p<0.05$. Values marked with * and ** are significant at $p<0.01$ and $p<0.001$, respectively.}
\begin{center}
  \resizebox{16cm}{!} {
    \begin{tabular}{l l l l l l l l l l l}
    \hline
                & WT                  & N348A              & N348W              & N348K              & Y211A             & Y211D              & Y211T              & vR111A             & N348A/Y211A        & N348W/Y211A        \\ \hline
    N348A       &  27.33              &                    &                    &                    &                   &                    &                    &                    &                    &                    \\
    N348W       & -39.87              & \textbf{-67.20}    &                    &                    &                   &                    &                    &                    &                    &                    \\
    N348K       & -40.64              & \textbf{-67.97}    & -0.77              &                    &                   &                    &                    &                    &                    &                    \\
    Y211A       & \textbf{-87.97*}    & \textbf{-115.30**} & -48.10             & -47.33             &                   &                    &                    &                    &                    &                    \\
    Y211D       & \textbf{103.44**}   & \textbf{76.12*}    & \textbf{143.31**}  & \textbf{144.08**}  & \textbf{191.42**} &                    &                    &                    &                    &                    \\
    Y211T       & \textbf{89.17*}     & \textbf{61.85}     & \textbf{129.04**}  & \textbf{129.81**}  & \textbf{177.15**} & -14.27             &                    &                    &                    &                    \\
   vR111A       & -17.76              & -44.94             & 22.26              & 23.03              & \textbf{70.36}    & \textbf{-121.05**} & \textbf{-106.78**} &                    &                    &                    \\
    N348A/Y211A & -9.04               & -36.37             & 30.83              & 31.60              & \textbf{78.93}    & \textbf{-112.48**} & \textbf{-98.21*}   & 8.57               &                    &                    \\
    N348W/Y211A & \textbf{-154.79**}  & \textbf{-182.12**} & \textbf{-114.92**} & \textbf{-114.15**} & \textbf{-66.2}    & \textbf{-258.23**} & \textbf{-243.96**} & \textbf{-137.18**} & \textbf{-145.75**} &                    \\
   vR111A/Y211A &  8.71               & -18.12             & 48.58              & 49.35              & \textbf{96.68*}   & \textbf{-94.74*}   & \textbf{-80.46*}   & 26.31              & 0.59               & \textbf{163.50**} \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

\clearpage
\begin {table}[H]
\caption{\label{tab:MAF_p}Pairwise difference $p-$values for maximum applied force. Bolded values are statistically significant at $p<0.05$.}
\begin{center}
  \resizebox{16cm}{!} {
    \begin{tabular}{l l l l l l l l l l l}
    \hline
                & WT                      & N348A                   & N348W                   & N348K                   & Y211A                   & Y211D                     & Y211T                   & vR111A                  & N348A/Y211A             & N348W/Y211A \\ \hline
    N348A       & 0.35                    &                         &                         &                         &                         &                           &                         &                         &                         &        \\
    N348W       & 0.22                    & \textbf{0.012}          &                         &                         &                         &                           &                         &                         &                         &        \\
    N348K       & 0.22                    & \textbf{0.012}          & 0.98                    &                         &                         &                           &                         &                         &                         &        \\
    Y211A       & \textbf{0.0045}         & \textbf{1.6x10$^{-05}$} & 0.14                    & 0.15                    &                         &                           &                         &                         &                         &        \\
    Y211D       & \textbf{0.00083}        & \textbf{0.0045}         & \textbf{3.9x10$^{-06}$} & \textbf{3.9x10$^{-06}$} & \textbf{5.7x10$^{-10}$} &                           &                         &                         &                         &        \\
    Y211T       & \textbf{0.0042}         & \textbf{0.022}          & \textbf{3.0x10$^{-05}$} & \textbf{3.0x10$^{-05}$} & \textbf{1.0x10$^{-08}$} & 0.67                      &                         &                         &                         &        \\
   vR111A       & 0.59                    & 0.11                    & 0.51                    & 0.51                    & \textbf{0.024}          & \textbf{8.9x10$^{-05}$}   & \textbf{0.00056}        &                         &                         &        \\
    N348A/Y211A & 0.78                    & 0.20                    & 0.35                    & 0.35                    & \textbf{0.011}          & \textbf{0.00026}          & \textbf{0.0016}         & 0.78                    &                         &        \\
    N348W/Y211A & \textbf{6.2x10$^{-07}$} & \textbf{9.3x10$^{-12}$} & \textbf{0.00021}        & \textbf{0.00026}        & \textbf{0.032}          & \textbf{$<$ 2x10$^{-16}$} & \textbf{2.4x10$^{-15}$} & \textbf{9.1x10$^{-06}$} & \textbf{2.8x10$^{-06}$} &        \\
   vR111A/Y211A & 0.78                    & 0.52                    & 0.14                    & 0.14                    & \textbf{0.0019}         & \textbf{0.0024}           & \textbf{0.010}          & 0.44                    & 0.59                    & \textbf{1.6x10$^{-07}$}  \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

\begin {table}[H]
\caption{\label{tab:AUC_p}Pairwise difference $p-$values for interpolated AUC. Bolded values are statistically significant at $p<0.05$.}
\begin{center}
  \resizebox{16cm}{!} {
    \begin{tabular}{l l l l l l l l l l l}
    \hline
                & WT               & N348A                   & N348W           & N348K                   & Y211A                   & Y211D                   & Y211T                   & vR111A          & N348A/Y211A & N348W/Y211A \\ \hline
    N348A       & 0.77             &                         &                 &                         &                         &                         &                         &                 &             &        \\
    N348W       & 0.67             & 0.42                    &                 &                         &                         &                         &                         &                 &             &        \\
    N348K       & 0.17             & 0.056                   & 0.38            &                         &                         &                         &                         &                 &             &        \\
    Y211A       & \textbf{0.00094} & \textbf{3.5x10$^{-05}$} & \textbf{0.0045} & 0.063                   &                         &                         &                         &                 &             &        \\
    Y211D       & \textbf{0.0072}  & \textbf{0.0051}         & \textbf{0.0016} & \textbf{4.6x10$^{-05}$} & \textbf{1.2x10$^{-09}$} &                         &                         &                 &             &        \\
    Y211T       & \textbf{0.016}   & \textbf{0.014}          & \textbf{0.0041} & \textbf{0.00016}        & \textbf{6.1x10$^{-09}$} & 0.77                    &                         &                 &             &        \\
   vR111A       & 0.38             & 0.18                    & 0.68            & 0.67                    & \textbf{0.016}          & \textbf{0.00032}        & \textbf{0.00094}        &                 &             &        \\
    N348A/Y211A & 0.061            & \textbf{0.013}          & 0.18            & 0.68                    & 0.17                    & \textbf{5.2x10$^{-06}$} & \textbf{2.1$^{-05}$}    & 0.38            &             &        \\
    N348W/Y211A & \textbf{0.00017} & \textbf{3.9x10$^{-06}$} & \textbf{0.0011} & \textbf{0.021}          & 0.68                    & \textbf{1.6x10$^{-10}$} & \textbf{6.3x10$^{-10}$} & \textbf{0.0047} & 0.063       &        \\
   vR111A/Y211A & 0.56             & 0.34                    & 0.86            & 0.47                    & \textbf{0.0072}         & \textbf{0.0010}         & \textbf{0.0025}         & 0.77            & 0.24        & \textbf{0.0019}  \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

>>>>

The statistical package R was used for all further analysis and visualization. Each of the 50 independent trajectories per mutant produced a fairly noisy force curve. The force curves for each mutant were smoothed either by using the smooth.spline and predict functions in R with default settings over all replicates as in Figure~\ref{fig:force_curve}. The latter approach was used for statistical analysis. The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC). We used the pairwise.t.test function with the FDR p-value correction to test for significance. The ggplot \citep{ggplot} package was used to generate most of the figures.

@@

We proceeded to compare our SMD results to that of the standard dual topology FEP approach to calculate relative free energy differences. The correlation between the energetically rigorous FEP and our statistical approach is high. For all 11 complexes tested, the correlation between max force and FEP was $r=-0.795$ at $p=0.0034$ (Figure~\ref{fig:fep}), and the correlation between AUC and FEP was $r=-0.593$ at $p=0.055$. Because of this strong correlation, we refer exclusively to the SMD results for the remainder of this work.

@@

\clearpage
\begin{figure}[p]
\centerline{\includegraphics[width=6in]{figures/fep_mutants.pdf}}
\caption{\label{fig:fep}\title{Max force versus free energy perturbation.} Scatter plot of maximum force in SMD versus the relative free energy difference calculated by FEP for all 10 mutants tested plus the WT complex. The WT complex for FEP was simply set to 0.0. The correlation between the two is $r=-0.795$ with $p=0.0034$.}
\end{figure}

@@

\clearpage
\begin {table}[H]
\caption{\label{tab:summary_mutants}Summary of prior information available for each mutation tested. Observed $in~vivo$ refers to mutations that have been observed in rodent populations. Phenotype $in~vitro$ refers to the observed phenotype in $in~vitro$ viral entry assays.} 
\begin{center}
  \resizebox{11cm}{!} {
    \begin{tabular}{l c r r}
    \hline
      Mutation & Observed $in~vivo$ & Phenotype $in~vitro$ \\ \hline
            WT &                Yes &         Normal Entry \\
         N348A &                 No &                    - \\
         N348K &                Yes &     Diminished Entry \\
         N348W &                 No &                    - \\
        vR111A &                 No &     Diminished Entry \\
   N348A/Y211A &                 No &                    - \\
  vR111A/Y211A &                 No &                    - \\
         Y211D &                Yes &        No Expression \\
         Y211T &                 No &     Diminished Entry \\
         Y211A &                 No &        No Expression \\
   N348W/Y211A &                 No &                    - \\

    \hline
    \end{tabular}
  }
\end{center}
\end{table}

\clearpage
\begin {table}[H]
\caption{\label{tab:summary_statistics}Summary statistics for each mutation tested. $\mu_\text{MAF}$ is the mean in piconewtons and $\sigma_\text{MAF}$ is the standard deviation of maximum applied force over all simulations. $\mu_\text{AUC}$ is the mean and $\sigma_\text{AUC}$ is the standard deviation of AUC over all simulations. $\Delta G$ is the free energy difference in kcal/mol calculated via FEP by the dual topology paradigm.}
\begin{center}
  \resizebox{15cm}{!} {
    \begin{tabular}{l c c c c c}
    \hline
      Mutation & $\mu_\text{MAF}$ (pN) & $\sigma_\text{MAF}$ & $\mu_\text{AUC}$ & $\sigma_\text{AUC}$ & $\Delta G$ (kcal/mol)\\ \hline
            WT &                734.4856 &              131.6513 &         145460.4 &            60232.26 &  0.000\\
         N348A &                748.5217 &              137.4864 &         133913.9 &            51078.64 & -2.149\\
         N348K &                705.0707 &              108.5079 &         141084.4 &            54450.28 & +3.184\\
         N348W &                697.3642 &              132.6436 &         136886.0 &            53796.44 & +3.033\\
        vR111A &                713.8081 &              106.7374 &         136103.2 &            52070.85 & +0.466\\ 
   N348A/Y211A &                703.7027 &              128.5866 &         113464.2 &            57451.62 & +5.203\\ 
  vR111A/Y211A &                741.0642 &              131.6287 &         130070.6 &            47665.56 & -2.440\\
         Y211D &                825.2586 &              115.4343 &         158878.7 &            63039.08 & +2.760\\
         Y211T &                806.8593 &              136.5648 &         167110.7 &            78849.29 & +0.875\\
         Y211A &                654.1138 &              108.5343 &         108090.0 &            43661.09 & +2.526\\
   N348W/Y211A &                594.9044 &              134.8233 &         108984.2 &            45451.00 & +8.206\\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

\clearpage
\begin {table}[H]
\caption{\label{tab:MAF_pairwise}Pairwise differences (row variable minus column variable) in mean maximum applied force. Bolded values are statistically significant at $p<0.05$.}
\begin{center}
  \resizebox{16cm}{!} {
    \begin{tabular}{l l l l l l l l l l l}
    \hline
             & WT                      & N348A                   & N348W                   & N348K                   & vR111A                  & N348A/Y211A               & vR111A/Y211A            & Y211D                   & Y211T                   & Y211A  \\ \hline
    N348A       & +14.036              &                         &                         &                         &                         &                           &                         &                         &                         &        \\
    N348W       & -29.414              & -43.451                 &                         &                         &                         &                           &                         &                         &                         &        \\
    N348K       & -37.121              & \textbf{-51.157}        & -7.7060                 &                         &                         &                           &                         &                         &                         &        \\
   vR111A       & -20.677              & -34.713                 & +8.7370                 & +16.443                 &                         &                           &                         &                         &                         &        \\
    N348A/Y211A & -30.782              & -44.819                 & -1.3670                 & +6.3380                 & -10.105                 &                           &                         &                         &                         &        \\
   vR111A/Y211A & +6.5790              & -7.4570                 & +35.993                 & +43.700                 & +27.256                 & +37.361                   &                         &                         &                         &        \\
    Y211D       & \textbf{+90.772}     & \textbf{+76.736}        & \textbf{+120.19}        & \textbf{+127.89}        & \textbf{+111.45}        & \textbf{+121.56}          & \textbf{+84.194}        &                         &                         &        \\
    Y211T       & \textbf{+72.373}     & \textbf{+58.337}        & \textbf{+101.79}        & \textbf{+109.50}        & \textbf{+93.051}        & \textbf{+103.16}          & \textbf{+65.795}        & -18.399                 &                         &        \\
    Y211A       & \textbf{-80.371}     & \textbf{-94.407}        & -50.956                 & -43.250                 & \textbf{-59.694}        & -49.588                   & \textbf{-86.950}        & \textbf{-171.1 4}       & \textbf{-152.75}        &        \\
    N348W/Y211A & \textbf{-139.58}     & \textbf{-153.62}        & \textbf{-110.17}        & \textbf{+102.46}        & \textbf{-118.903}       & \textbf{-108.80}          & \textbf{+146.16}        & \textbf{+230.35}        & \textbf{-211.95}        & \textbf{-59.209}  \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

\clearpage
\begin {table}[H]
\caption{\label{tab:MAF_p}Pairwise difference $p-$values for maximum applied force. Bolded values are statistically significant at $p<0.05$.}
\begin{center}
  \resizebox{16cm}{!} {
    \begin{tabular}{l l l l l l l l l l l}
    \hline
                & WT                      & N348A                   & N348W                   & N348K                   & vR111A                  & N348A/Y211A               & vR111A/Y211A            & Y211D                   & Y211T                   & Y211A  \\ \hline
    N348A       & 0.60                    &                         &                         &                         &                         &                           &                         &                         &                         &        \\
    N348W       & 0.31                    & 0.077                   &                         &                         &                         &                           &                         &                         &                         &        \\
    N348K       & 0.20                    & \textbf{0.038}          & 0.81                    &                         &                         &                           &                         &                         &                         &        \\
   vR111A       & 0.51                    & 0.16                    & 0.79                    & 0.60                    &                         &                           &                         &                         &                         &        \\
    N348A/Y211A & 0.29                    & 0.07                    & 0.95                    & 0.81                    & 0.77                    &                           &                         &                         &                         &        \\
   vR111A/Y211A & 0.82                    & 0.79                    & 0.21                    & 0.13                    & 0.35                    & 0.20                      &                         &                         &                         &        \\
    Y211D       & \textbf{0.00093}        & \textbf{0.0012}         & \textbf{1.4x10$^{-5}$}  & \textbf{5.0x10$^{-6}$}  & \textbf{5.6x10$^{-5}$}  & \textbf{1.2x10$^{-5}$}    & \textbf{0.0022}         &                         &                         &        \\
    Y211T       & \textbf{0.01}           & \textbf{0.018}          & \textbf{0.00022}        & \textbf{8.7x10$^{-5}$}  & \textbf{0.0008}         & \textbf{0.0002}           & \textbf{0.021}          & 0.56                    &                         &        \\
    Y211A       & \textbf{0.0034}         & \textbf{7.2x10$^{-05}$} & 0.074                   & 0.13                    & \textbf{0.035}          & 0.079                     & \textbf{0.0016}         & \textbf{4.2x10$^{-10}$} & \textbf{4.2x10$^{-8}$}  &        \\
    N348W/Y211A & \textbf{3.9x10$^{-7}$}  & \textbf{1.1x10$^{-10}$} & \textbf{6.5x10$^{-5}$}  & \textbf{0.00021}        & \textbf{1.6x10$^{-5}$}  & \textbf{7.2x10$^{-5}$}    & \textbf{1.3x10$^{-7}$}  & \textbf{< 2x10$^{-16}$} & \textbf{2.0x10$^{-14}$} & \textbf{0.036}  \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

\begin {table}[H]
\caption{\label{tab:AUC_p}Pairwise difference $p-$values for interpolated AUC. Bolded values are statistically significant at $p<0.05$.}
\begin{center}
  \resizebox{16cm}{!} {
    \begin{tabular}{l l l l l l l l l l l}
    \hline
                & WT                      & N348A                   & N348W                   & N348K                   & vR111A                  & N348A/Y211A               & vR111A/Y211A            & Y211D                   & Y211T                   & Y211A  \\ \hline
    N348A       & 0.33                    &                         &                         &                         &                         &                           &                         &                         &                         &        \\
    N348W       & 0.76                    & 0.59                    &                         &                         &                         &                           &                         &                         &                         &        \\
    N348K       & 0.59                    & 0.80                    & 0.76                    &                         &                         &                           &                         &                         &                         &        \\
   vR111A       & 0.55                    & 0.85                    & 0.76                    & 0.94                    &                         &                           &                         &                         &                         &        \\
    N348A/Y211A & \textbf{0.017}          & 0.07                    & \textbf{0.031}          & 0.076                   & 0.08                    &                           &                         &                         &                         &        \\
   vR111A/Y211A & 0.26                    & 0.76                    & 0.46                    & 0.68                    & 0.72                    & 0.22                      &                         &                         &                         &        \\
    Y211D       & 0.33                    & \textbf{0.029}          & 0.18                    & 0.09                    & 0.08                    & \textbf{0.00046}          & \textbf{0.029}          &                         &                         &        \\
    Y211T       & 0.09                    & \textbf{0.0056}         & \textbf{0.046}          & \textbf{0.027}          & \textbf{0.023}          & \textbf{4.1x10$^{-5}$}    & \textbf{0.006}          & 0.59                    &                         &        \\
    Y211A       & \textbf{0.0056}         & \textbf{0.027}          & \textbf{0.016}          & \textbf{0.029}          & \textbf{0.031}          & 0.75                      & 0.09                    & \textbf{8.2x10$^{-5}$}  & \textbf{8.5x10$^{-6}$}  &        \\
    N348W/Y211A & \textbf{0.006}          & \textbf{0.029}          & \textbf{0.017}          & \textbf{0.032}          & \textbf{0.034}          & 0.76                      & 0.1                     & \textbf{9.4x10$^{-5}$}  & \textbf{8.5x10$^{-6}$}  & 0.94  \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}
